---
title: "Activity Analysis"
author: "Yashu Seth"
date: "May 11, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('ggplot2')
```

- **Code for reading in the dataset and/or processing the data**
```{r read_data}
setwd("C:/work/R")
data <- read.csv("activity.csv", colClasses = 'character')
```

- **Histogram of the total number of steps taken each day**
```{r hist_total_steps}
tot_steps <- tapply(as.integer(data$steps), data$date, sum, na.rm=T)
x <- strptime(rownames(tot_steps), "%Y-%m-%d")
y <- as.vector(tot_steps)
qplot(x,y) + geom_bar(stat = 'identity') + xlab("Date") + 
    ylab("Total number of Steps")
```

- **Mean and median number of steps taken each day**
```{r mean_median}
mean_steps <- tapply(as.integer(data$steps), data$date, mean,
                     na.rm=T)
```
The mean number of steps taken in each day -
```{r}
mean_steps

median_steps <- tapply(as.integer(data$steps), data$date, median,
                     na.rm=T)
```
The median number of steps taken in each day -
```{r}
median_steps
```

- **Time series plot of the average number of steps taken**
```{r avg_plot}
y <- as.vector(mean_steps)
qplot(x,y) + geom_line() + xlab("Date") + ylab("Average number of Steps")
```

- **The 5-minute interval that, on average, contains the maximum number of steps**
```{r 5_min_average}
int_avg <- tapply(as.integer(data$steps),
                  data$interval, mean, na.rm = T)
```


Therefore the 5-minute interval that, on average, contains the maximum number of steps is -
```{r}
rownames(int_avg)[int_avg==max(int_avg)]
```

- **Code to describe and show a strategy for imputing missing data**
I will replace the missing values in the steps column with the mean steps of that particular day. But there are certain missing values in the mean steps data as well. I will replace this with the total mean.
```{r}
total_mean <- mean(mean_steps, na.rm = T)
mean_steps[is.nan(mean_steps)] <- total_mean
```
Now I will replace the missing values in the steps columns with the average number of steps taken that day.
```{r}
for(i in 1:nrow(data)){
    row <- data[i,]
    if(is.na(row$steps)){
        row$steps <- mean_steps[row$date]
    }
    data[i,] <- row
}
```

- **Histogram of the total number of steps taken each day after missing values are imputed**
```{r hist_total_steps_aft_missing}
tot_steps <- tapply(as.integer(data$steps), data$date, sum, na.rm=T)
x <- strptime(rownames(tot_steps), "%Y-%m-%d")
y <- as.vector(tot_steps)
qplot(x,y) + geom_bar(stat = 'identity') + xlab("Date") + 
    ylab("Total number of Steps")
```

- **Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends**
```{r}
day <- weekdays(strptime(data$date, "%Y-%m-%d"))
r <- day=="Saturday" | day=="Sunday"
day[r] <- "weekend"
day[!r] <- "weekday"
data <- cbind(data, day)
day_data <- split(data, data$day)
int_avg_wd <- tapply(as.integer(day_data$weekday$steps),
                  day_data$weekday$interval, mean, na.rm = T)
int_avg_we <- tapply(as.integer(day_data$weekend$steps),
                  day_data$weekend$interval, mean, na.rm = T)
par(mfrow=c(2,1))
plot(rownames(int_avg_wd), as.vector(int_avg_wd), pch=20, 
     xlab='Intervals on weekdays', ylab = 'average steps', type = 'l')
plot(rownames(int_avg_we), as.vector(int_avg_we), pch=20, type = 'l',
     xlab='Intervals on weekends', ylab = 'average steps')
```

- **All of the R code needed to reproduce the results (numbers, plots, etc.) in the report**

```{r , eval=FALSE}
library('ggplot2')

setwd("C:/work/R")
data <- read.csv("activity.csv", colClasses = 'character')

tot_steps <- tapply(as.integer(data$steps), data$date, sum, na.rm=T)
x <- strptime(rownames(tot_steps), "%Y-%m-%d")
y <- as.vector(tot_steps)
qplot(x,y) + geom_bar(stat = 'identity') + xlab("Date") + 
    ylab("Total number of Steps")

mean_steps <- tapply(as.integer(data$steps), data$date, mean,
                     na.rm=T)

mean_steps

median_steps <- tapply(as.integer(data$steps), data$date, median,
                     na.rm=T)

median_steps

y <- as.vector(mean_steps)
qplot(x,y) + geom_line() + xlab("Date") + ylab("Average number of Steps")

int_avg <- tapply(as.integer(data$steps),
                  data$interval, mean, na.rm = T)

rownames(int_avg)[int_avg==max(int_avg)]

total_mean <- mean(mean_steps, na.rm = T)
mean_steps[is.nan(mean_steps)] <- total_mean

for(i in 1:nrow(data)){
    row <- data[i,]
    if(is.na(row$steps)){
        row$steps <- mean_steps[row$date]
    }
    data[i,] <- row
    
tot_steps <- tapply(as.integer(data$steps), data$date, sum, na.rm=T)
x <- strptime(rownames(tot_steps), "%Y-%m-%d")
y <- as.vector(tot_steps)
qplot(x,y) + geom_bar(stat = 'identity') + xlab("Date") + 
    ylab("Total number of Steps")

day <- weekdays(strptime(data$date, "%Y-%m-%d"))
r <- day=="Saturday" | day=="Sunday"
day[r] <- "weekend"
day[!r] <- "weekday"
data <- cbind(data, day)
day_data <- split(data, data$day)
int_avg_wd <- tapply(as.integer(day_data$weekday$steps),
                  day_data$weekday$interval, mean, na.rm = T)
int_avg_we <- tapply(as.integer(day_data$weekend$steps),
                  day_data$weekend$interval, mean, na.rm = T)
par(mfrow=c(2,1))
plot(rownames(int_avg_wd), as.vector(int_avg_wd), pch=20, 
     xlab='Intervals on weekdays', ylab = 'average steps', type = 'l')
plot(rownames(int_avg_we), as.vector(int_avg_we), pch=20, type = 'l',
     xlab='Intervals on weekends', ylab = 'average steps')
```

